---
title: "WIDS-modeling"
author: "Math 4180"
date: "2/28/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Set Up

## Libraries
You will need to install any of these packages if you do not already have them. 
```{r load-libraries, warning=F, message=F}
library(tidyverse)
library(data.table) #for reading in big data from a url
library(janitor) #for tably function (similar to table function)
library(corrr) # for correlations
library(skimr) # for skim function (similar to summary function)
library(tidymodels) #for modeling
library(themis) #for downsampling
```


```{r load-data, warning=F, message=F}
training <- fread("https://math4180.netlify.app/data/TrainingWiDS2021.csv")
dictionary <- fread("https://math4180.netlify.app/data/DataDictionaryWiDS2021.csv")

training %>% skimr::skim()
```

```{r cleaning-data, eval=F, warning=F, message=F}
diabetes <- training %>%
  select(everything(), -contains("h1"), -V1, -encounter_id) %>%
  mutate(gender = replace_na(gender, "Unknown")) %>%
  mutate(ethnicity = replace_na(ethnicity, "Other/Unknown")) %>%
  mutate(hospital_admit_source = replace_na(hospital_admit_source, "Unknown")) %>%
  mutate(icu_admit_source = replace_na(icu_admit_source, "Unknown")) %>%
  mutate_if(is.character, factor) %>%
  mutate(diabetes_mellitus = as.factor(diabetes_mellitus))
```

Feature Engineering

```{r}
diabetes <- training %>%
  select(everything(), -contains("h1"), -V1, -encounter_id) %>%
  group_by(gender) %>%
  mutate(height = coalesce(height, mean(height, na.rm = TRUE))) %>%
  mutate(weight = coalesce(weight, mean(weight, na.rm = TRUE))) %>%
  ungroup() %>%
  mutate(bmi = (weight/(height/100)^2))%>%
  mutate(gender = replace_na(gender, "Unknown")) %>%
  mutate(ethnicity = replace_na(ethnicity, "Other/Unknown")) %>%
  mutate(diabetes_mellitus = factor(diabetes_mellitus))
```

```{r, eval=F}
diabetes <- diabetes %>% 
  mutate(bmi_cats = case_when(bmi < 18.5 ~ "Underweight",
                              18.5 <= bmi & bmi <25 ~ "Normal",
                              25 <= bmi & bmi < 30 ~ "Overweight",
                              30 <= bmi & bmi < 35 ~ "Class I Obesity",
                              35 <= bmi & bmi < 40 ~ "Class II Obesity",
                              40 <= bmi  ~ "Class III Obesity",
                              TRUE ~ "Unknown")) %>%
  mutate(age_cats = case_when(15 <= age & age <25  ~ "Youth",
                              25 <= age & age < 65 ~ "Adult",
                              age >= 65 ~ "Senior",
                              TRUE ~ "Unknown")) %>%
  mutate(across(contains("glucose"), 
                ~ case_when( . < 60 ~ "Hypoglycemia",
                             60 <= . & . < 70 ~ "Early Hypoglycemia",
                             70 <= . & . <140 ~ "Normal",
                             140 <= . & . < 200 ~ "Early Diabetic",
                             . > 200 ~ "Diabetic",
                             TRUE ~ "Unknown"),
                .names = "{col}_cats")) %>%
  mutate(across(contains("sys"), 
                ~ case_when( . < 120 ~ "Normal",
                             120<= . & . < 130 ~ "Elevated",
                             130 <= . & . < 140 ~ "Hypertension Stage 1",
                             140 <= . & . < 180 ~ "Hypertension Stage 2",
                             180 <= . ~ "Hypertensive Crisis",
                             TRUE ~ "Unknown"),
                .names = "{col}_cats")) %>%
  mutate(across(contains("diasbp"), 
                ~ case_when( . < 80 ~ "Normal",
                             80 <= . & . < 90~ "Hypertension Stage 1",
                             90 <= . & . < 120 ~ "Hypertension Stage 2",
                             180 <= . ~ "Hypertensive Crisis",
                             TRUE ~ "Unknown"),
                .names = "{col}_cats")) %>%
  mutate(across(contains("creat"), 
                ~ case_when( gender == "M" & . >= 0.6 & . <= 1.2 ~ "Normal",
                             gender == "F" & . >= 0.5 & . <= 1.1 ~ "Normal",
                             is.numeric(.) ~ "Abnormal",
                             TRUE ~ "Unknown"),
                .names = "{col}_cats")) 
```

# Modeling

## Spliting data
Set up data for model building and evaluation 
```{r split-by-strata}
#split data randomly but use equal proportions of target variable
set.seed(2021)
diabetes_split <- initial_split(diabetes, strata = diabetes_mellitus)

#create new training and evaluation sets for modeling
diabetes_train <- training(diabetes_split)
diabetes_eval <- testing(diabetes_split)
```

Check that the strata are appropriate
```{r strata-distribution}
#original distribution of target variable (strata variable)
janitor::tabyl(training$diabetes_mellitus)

#new training set strata
dim(diabetes_train)
janitor::tabyl(diabetes_train$diabetes_mellitus)

#new evaluation set strata
dim(diabetes_eval)
janitor::tabyl(diabetes_eval$diabetes_mellitus)

```


## Recipe
Define recipe for modeling and apply to our new datasets
```{r recipe_and_bake}
diabetes_recipe <- recipe(diabetes_mellitus ~., data = diabetes_train) %>%
  step_nzv(all_numeric(), -all_outcomes()) %>%
  #step_rm(bmi, contains("glucose"), contains("diasp"), contains("sys"), contains("creatin")) %>%
  step_knnimpute(age) %>%
  step_other(all_nominal(), -all_outcomes(), threshold = 0.01) %>%
  step_dummy(all_nominal(), -all_outcomes(), one_hot = TRUE) %>%
  step_corr(all_numeric(), -all_outcomes(), threshold = 0.5) %>%
  step_normalize(all_numeric()) %>%
  step_mutate_at(all_numeric(), fn=~ if_else(is.na(.), -999, . )) %>%
  prep()
#apply recipe to training set
train_baked <- bake(diabetes_recipe, new_data = diabetes_train)

#apply recipe to evaluation set
eval_baked <- bake(diabetes_recipe, new_data = diabetes_eval)


```

## Model Selection
Declaring model selection
```{r model-selection}
# Specify the model: set the engine (ie what library you're using) and set the mode (regression or classification)
log_spec <- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification")

null_spec <- null_model() %>%
  set_engine("parsnip") %>%
  set_mode("classification")

tree_spec <- decision_tree(tree_depth = 10) %>%
  set_engine("rpart") %>%
  set_mode("classification")
```

Fitting to our training set
```{r model-fit}
#fit model specification to our training data where our target variable is diabetes_mellitus
log_fit <- log_spec %>%
  fit(diabetes_mellitus~.,
      data = train_baked)

null_fit <- null_spec %>%
  fit(diabetes_mellitus~.,
      data = train_baked)

tree_fit <- tree_spec %>%
  fit(diabetes_mellitus~.,
      data = train_baked)
```

Evaluation using the evaluation set
```{r evaluation}
#using the fitted data, predict the target variable for the evaluation set
log_preds<- log_fit %>%
  predict(eval_baked) %>%
  bind_cols(eval_baked %>% select(diabetes_mellitus))

null_preds<- null_fit %>%
  predict(eval_baked) %>%
  bind_cols(eval_baked %>% select(diabetes_mellitus))

tree_preds<- tree_fit %>%
  predict(eval_baked) %>%
  bind_cols(eval_baked %>% select(diabetes_mellitus))
```

Visualizing predictions and outputting accuracy
```{r eval-viz}
#visualize the accuracy using a confusion matrix
log_preds %>%
  conf_mat(diabetes_mellitus, .pred_class) %>%
  pluck(1) %>%
  as_tibble() %>%
  ggplot(aes(Prediction, Truth, alpha = n)) +
  geom_tile(show.legend = FALSE) +
  geom_text(aes(label = n), colour = "white", alpha = 1, size = 8)+
  labs(title = "Logistic Predictions")

null_preds %>%
  conf_mat(diabetes_mellitus, .pred_class) %>%
  pluck(1) %>%
  as_tibble() %>%
  ggplot(aes(Prediction, Truth, alpha = n)) +
  geom_tile(show.legend = FALSE) +
  geom_text(aes(label = n), colour = "white", alpha = 1, size = 8)+
  labs(title = "Null Prediction")

tree_preds %>%
  conf_mat(diabetes_mellitus, .pred_class) %>%
  pluck(1) %>%
  as_tibble() %>%
  ggplot(aes(Prediction, Truth, alpha = n)) +
  geom_tile(show.legend = FALSE) +
  geom_text(aes(label = n), colour = "white", alpha = 1, size = 8)+
  labs(title = "Tree Prediction")
  
#output accuracy
log_preds %>%
  metrics(diabetes_mellitus, .pred_class) %>%
  select(-.estimator) %>%
  filter(.metric == "accuracy") 

null_preds %>%
  metrics(diabetes_mellitus, .pred_class) %>%
  select(-.estimator) %>%
  filter(.metric == "accuracy") 

tree_preds %>%
  metrics(diabetes_mellitus, .pred_class) %>%
  select(-.estimator) %>%
  filter(.metric == "accuracy") 
```

# Next Steps

- How should we handle missing numeric values? (i.e. our vitals)

- How does this model compare to a null model? How could we implement this? (hint: you can create a null model using the same set up as the logistic model we did)

- What other models would be useful given that we have a binary output?

- Should we include other variables?